Trajectory Mapping Results

### 2020-12-16 workflow/scripts/analyze_trajectories_v2.Rmd
wip analyze traj Rmd

case_data <- get_cases(demes, from = ymd("2019-09-01"), to = MRS + 10) %>%
  arrange(date) %>%
  group_by(deme) %>%
  mutate(cumcases = cumsum(cases),
         cumdeaths = cumsum(deaths)) %>%
  group_by(deme, date) %>%
  mutate(cumcases = max(cumcases),
         cumdeaths = max(cumdeaths)) %>%
  select(deme, date, cumcases, cumdeaths, cases)  %>%
  distinct()

@@

```{r trajsv2}
# 1.2 All trajectories single deme in log scale with ECDC case count data
trajs <- lapply(demes$deme, function(deme) {
  gt_v2 %>%
    # 1. Prepare data
    filter(type == deme) %>%
    mutate(traj = as.character(traj)) %>%
    bind_rows(case_data %>%
                rename(type = deme,
                       cumNt = cumcases) %>%
                filter(type == deme) %>%
                mutate(traj = "ecdc") %>%
                select(type, traj, date, cumNt) %>%
                distinct()) %>%
    # 2. Plot
    ggplot() +
    geom_line(aes(date, cumNt, group = traj,
                  color = traj == "ecdc",
                  alpha = traj == "ecdc",
                  linetype = traj == "ecdc",
                  size = traj == "ecdc")) +
    scale_color_manual(name = "",
                       labels = c(paste0(deme, " inferred\n population trajectories"),
                                  paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                         " ECDC total case count")),
                       values = c(dcolors[[deme]], dark_dcolors[[deme]])) +
    scale_alpha_manual(name = "", values = c(0.2, 1), guide = FALSE) +
    scale_linetype_manual(name = "", values = c(1, 2),
                          labels = c(paste0(deme, " inferred\n population trajectories"),
                                     paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                            " ECDC total case count"))) +
    scale_size_manual(name = "", values = c(0.5, 1),
                      labels = c(paste0(deme, " inferred\n population trajectories"),
                                 paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                        " ECDC total case count"))) +
    scale_x_date(limits = c(ymd("2019-11-01"), ymd(MRS + 10)), date_breaks = "2 weeks", 
                 date_labels = "%b %d") +
    scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
    ylab("Population size") +
    theme(legend.position = c(0.4, 0.9),
          axis.title.x=element_blank())
  })

@@

```{r report-rate}
# 1.3. Reporting rate by day by case
reportrate <- gt_summary_v2 %>%
  filter(!type %in% c(china_deme, "OtherEuropean")) %>%
  left_join(case_data %>% rename(type = deme), by = c("type", "date"))  %>%
  replace_na(list(cases = 0)) %>%
  mutate(reportrate = ifelse(is.infinite(cases/Dmedian), NA, cases/Dmedian),
         rlow = ifelse(is.infinite(cases/Dhigh), NA, cases/Dhigh),
         rhigh = ifelse(is.infinite(cases/Dlow), NA, cases/Dlow)) %>%
  filter(!is.na(reportrate)) %>%
  # mutate(week = lubridate::week(date)) %>%
  # group_by(week, type) %>%
  # mutate(w_reportrate = mean(reportrate)) %>%
  ggplot() +
  geom_line(aes(x = date, y = reportrate, color = type), stat="identity") +
  geom_point(aes(x = date, y = reportrate, color = type), stat="identity") +
  geom_ribbon(aes(date, ymin = rlow, ymax = rhigh, fill = type), alpha = 0.1) +
  ylab("Reporting rate") +
  scale_color_manual(values = dcolors, name = "") +
  scale_fill_manual(values = dcolors, name = "") +
  scale_x_date(limits = c(ymd("2020-01-20"), ymd(MRS + 10)), date_breaks = "1 week", date_labels = "%b %d") +
  theme(#legend.position = "none",
        axis.title.x=element_blank()) +
  coord_cartesian(ylim = c(0,5))

@@

## Inferred case counts - Additional figures
```{r}
# 1.1 Median and 95% CCI by deme in log scale
gribbon <- ggplot(gt_summary) +
  geom_ribbon(aes(date, ymin = Ilow, ymax = Ihigh, fill = type), alpha = 0.5) +
  geom_line(aes(date, Imedian, colour = type)) +
  ylab("Population size") +
  xlab("Date") +
  #labs(title = "SARS-CoV-2 early epidemics in Europe and Hubei",
       #subtitle = "Case trajectories Median and 95% CCI by deme") +
  scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
  scale_x_date(limits = c(ymd("2019-10-01"), ymd("2020-03-08")), date_breaks = "1 month", date_labels = "%b %d") +
  scale_fill_manual(name = "", values = dcolors) +
  scale_color_manual(name = "", values = dcolors) +
  theme(legend.position = c(0.2, 0.7),
        legend.box="horizontal",
        axis.title.x=element_blank())
#
# ggexport(gribbon,
#          filename = paste0(args$output_figure, "01.png"),
#          width = 1500, height = 1000, res = 300)
```

>>>>

case_data <- get_cases(demes, from = ymd("2019-09-01"), to = MRS + 10) %>%
  arrange(date) %>%
  group_by(deme) %>%
  mutate(cumcases = cumsum(cases),
         cumdeaths = cumsum(deaths)) %>%
  group_by(deme, date) %>%
  mutate(cumcases = max(cumcases),
         cumdeaths = max(cumdeaths)) %>%
  select(deme, date, cumcases, cumdeaths, cases)  %>%
  distinct_at(vars(deme, date, cumcases, cumdeaths), .keep_all = TRUE)

@@

```{r trajsv2}
# 1.2 All trajectories single deme in log scale with ECDC case count data
trajs <- lapply(demes$deme, function(deme) {
  gt_v2 %>%
    # 1. Prepare data
    filter(type == deme) %>%
    mutate(traj = as.character(traj)) %>%
    bind_rows(case_data %>%
                rename(type = deme,
                       cumNt = cumcases) %>%
                filter(type == deme) %>%
                mutate(traj = "ecdc") %>%
                select(type, traj, date, cumNt) %>%
                distinct()) %>%
    # 2. Plot
    ggplot() +
    geom_line(aes(date, cumNt, group = traj,
                  color = traj == "ecdc",
                  alpha = traj == "ecdc",
                  linetype = traj == "ecdc",
                  size = traj == "ecdc")) +
    scale_color_manual(name = "",
                       labels = c(paste0(deme, " inferred\n population trajectories"),
                                  paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                         " ECDC total case count")),
                       values = c(dcolors[[deme]], dark_dcolors[[deme]])) +
    scale_alpha_manual(name = "", values = c(0.2, 1), guide = FALSE) +
    scale_linetype_manual(name = "", values = c(1, 2),
                          labels = c(paste0(deme, " inferred\n population trajectories"),
                                     paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                            " ECDC total case count"))) +
    scale_size_manual(name = "", values = c(0.5, 1),
                      labels = c(paste0(deme, " inferred\n population trajectories"),
                                 paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                        " ECDC total case count"))) +
    scale_x_date(limits = c(ymd("2019-11-01"), ymd(MRS + 10)), date_breaks = "2 weeks", 
                 date_labels = "%b %d") +
    scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
    ylab("Population size") +
    theme(legend.position = c(0.3, 0.9),
          axis.title.x=element_blank())
  })

@@

```{r report-rate, fig.height = 4}
# 1.3. Reporting rate by day by case
reportrate <- gt_summary_v2 %>%
  filter(!type %in% c(china_deme, "OtherEuropean")) %>%
  left_join(case_data %>% rename(type = deme), by = c("type", "date"))  %>%
  replace_na(list(cases = 0)) %>%
  mutate(reportrate = ifelse(is.infinite(cases/Dmedian), NA, cases/Dmedian),
         rlow = ifelse(is.infinite(cases/Dhigh), NA, cases/Dhigh),
         rhigh = ifelse(is.infinite(cases/Dlow), NA, cases/Dlow)) %>%
  filter(!is.na(reportrate)) %>%
  # mutate(week = lubridate::week(date)) %>%
  # group_by(week, type) %>%
  # mutate(w_reportrate = mean(reportrate)) %>%
  ggplot() +
  geom_line(aes(x = date, y = reportrate, color = type), stat="identity") +
  geom_point(aes(x = date, y = reportrate, color = type), stat="identity") +
  geom_ribbon(aes(date, ymin = rlow, ymax = rhigh, fill = type), alpha = 0.1) +
  ylab("Reporting rate") +
  scale_color_manual(values = dcolors, name = "") +
  scale_fill_manual(values = dcolors, name = "") +
  scale_x_date(limits = c(ymd("2020-01-20"), ymd(MRS + 10)), date_breaks = "1 week", date_labels = "%b %d") +
  theme(#legend.position = "none",
        axis.title.x=element_blank()) +
  coord_cartesian(ylim = c(0,5))

@@

## Inferred case counts - Additional figures
```{r traj-sum, fig.height = 4}
# 1.1 Median and 95% CCI by deme in log scale
gribbon <- ggplot(gt_summary) +
  geom_ribbon(aes(date, ymin = Ilow, ymax = Ihigh, fill = type), alpha = 0.5) +
  geom_line(aes(date, Imedian, colour = type)) +
  ylab("Population size") +
  xlab("Date") +
  #labs(title = "SARS-CoV-2 early epidemics in Europe and Hubei",
       #subtitle = "Case trajectories Median and 95% CCI by deme") +
  scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
  scale_x_date(limits = c(ymd("2019-10-01"), ymd("2020-03-08")), date_breaks = "1 month", date_labels = "%b %d") +
  scale_fill_manual(name = "", values = dcolors) +
  scale_color_manual(name = "", values = dcolors) +
  theme(legend.position = c(0.2, 0.7),
        legend.box="horizontal",
        axis.title.x=element_blank())

gribbon

#
# ggexport(gribbon,
#          filename = paste0(args$output_figure, "01.png"),
#          width = 1500, height = 1000, res = 300)
```
